### 2023-01-27 outputs/paper/paper.qmd
Download raw data and create paper

---
title: "Analysis of Immunization Coverage in Toronto"
author: "Julie Nguyen"
format: pdf
abstract: this is an abstract
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2023-01-27 scripts/01_download_data.R

@@ -1,0 +1,19 @@
+library(opendatatoronto)
+library(dplyr)
+
+# get package
+package <- show_package("99ff3657-b2e7-4005-a6fd-c36838ccc96d")
+package
+
+# get all resources for this package
+resources <- list_package_resources("99ff3657-b2e7-4005-a6fd-c36838ccc96d")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% 'csv')
+datastore_resources
+
+# load the immunization coverage in the AY 2017-2018
+immunization_2017_2018 <- filter(datastore_resources, row_number()==1) %>% get_resource()
+
+# load the immunization coverage in the AY 2018-2019
+immunization_2018_2019 <- filter(datastore_resources, row_number()==1) %>% get_resource()

### 2023-01-27 scripts/03_clean_data.R
clean and visualize data

@@ -1,0 +1,23 @@
+library(opendatatoronto)
+library(dplyr)
+library(janitor)
+
+outbreak_2020 %>% clean_names()
+outbreak_2021 %>% clean_names()
+outbreak_2022 %>% clean_names()
+
+ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak)
+ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak)
+ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak)
+
+ob_by_type <- merge(merge(ob_by_type_2020, ob_by_type_2021, by = "Type.of.Outbreak", all.x = TRUE), ob_by_type_2022, by = "Type.of.Outbreak", all.x = TRUE)
+
+ob_by_type_2 <- merge(merge(ob_by_type_2020, ob_by_type_2021, by = 0), ob_by_type_2022, by = 0)
+
+ob_by_type %>% rename("Type.of.Outbreak" = "Outbreak_Type")
+
+#ggplot(data=ob_by_type, aes(x=dose, y=len, group=supp)) + geom_line()+ geom_point()
+# Change line types
+ggplot(data=df2, aes(x=dose, y=len, group=supp)) +
+  geom_line(linetype="dashed", color="blue", size=1.2)+
+  geom_point(color="red", size=3)

### 2023-02-01 outputs/paper/paper.qmd
clean and generate graph for each task

@@
 ---
-title: "Analysis of Immunization Coverage in Toronto"
+title: "Outbreak Management in Toronto Health Institutions"
 author: "Julie Nguyen"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 format: pdf
-abstract: this is an abstract
+output:
+  bookdown::pdf_document2:
+abstract: Outbreak management has always been one of the most important mission for healthcare institutions, not only to prevent transmissions of diseases/illnesses but also to protect public health.
+Especially in the COVID-19 era, outbreak response and management is considered as the top priority in healthcare institution operation.
+This paper discusses the big picture of outbreak management in Toronto health institutions from 2020 to 2022, as well as pointing out the relationship between healthcare setting, outbreak type and outbreak response.
+execute: 
+  message: false
+  warning: false
 ---
 
-## Quarto
+## Introduction
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+  In the previous years, a lot has been learned with regards to the COVID-19 pandemic.
+With the unprecedented cases increase of different diseases, the responsibility for outbreak response and management in healthcare institutions has become more important than ever before.
+Outbreak management helps stop the further transmission of illnesses/diseases by identifying the source and implementing control measures.
 
-## Running Code
+  In order to protect and improve public health, the Toronto Public Health works with healthcare institutions to prevent and control institutional outbreaks, including the use of best practices in infection prevention control.
+Healthcare institutions are required to monitor staff and patients/residents for signs and symptoms of gastroenteric and respiratory infections under the Ontario Health Protection and Promotion Act (HPPA).
+(CITE) The institutions must also actively seek, detect and report suspected and/or confirmed outbreaks to their local public health unit.
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+  In the following paper, I will delve into the Outbreaks in Toronto Healthcare Institutions in the previous years to shed light on the differences and trends in reacting to outbreaks.
+I will first guide you through the data collection and analyzing process, which will be explained thoroughly in Section 2.
+Then I will visualize the data and explain the implications of the data, finally draw some limitations and conclusions of this data set.
+
+## Data
+### Data Collection
+
+  All the data used in this paper was obtained from the City of Toronto Open Data Portal and is titled "Outbreaks in Toronto Healthcare Institutions".
+The dataset is published by Toronto Public Health and is refreshed weekly, with the last refresh on Jan 19, 2023.
+The data provided on the portal include subsets of data from the year 2003.
+Nevertheless, I will just focus on the data from 2020 to 2022 in order to witness the changes in outbreak management during the COVID-19 era.
+According to opendatatoronto (cite), an outbreak is defined as "a localized increase (e.g.
+in an institution, or a specific ward or floor within an institution) in the rate of infection or illness, above that which is expected."
+
+  I first acquired the data using opendatatoronto package [@datatoronto] and imported to RStudio using the R package [cite].
+To process the data, all the necessary libraries into R script, including tidyverse [@tidyverse] and janitor [@janitor] to clean and manipulate the data, ggplot2 [@ggplot], knitr [@knitr], RColorBrewer [@RColor], and kableExtra [@kableExtra] to visualize the data.
+
+  After loading the data to the local environment, I started to create new dataframes.
+As the opendatatoronto @datatoronto package consists of multiple datasets by year, I had to download and extract the datasets of the year 2020, 2021, and 2022 separately, and  calling it *outbreak_2020*, *outbreak_2021*, and *outbreak_2022* respectively (ADD FOOTNOTE HERE).
+
+### Data Cleaning
+
+  For data cleaning, the target samples and focus were the type of outbreaks.
+I first observed the number occurrences of outbreaks from 2020 to 2022 in order to understand context of how outbreaks emerged.
+I also considered the type of outbreaks happened in each setting to understand whether the setting affect the control of outbreak.
+I thus filtered the datasets based on those criteria using tidyverse (@tidyverse) and janitor (@janitor), and dplyr (@dplyr).
+  I then created two dataframes by outbreak type and outbreak setting in order to witnesss
+
+### Analyzing and Visualizing the Data
+
+  The Toronto Health Institutions divided the outbreaks into 03 main types: Enteric, Respiratory, and other types.
+Throughout the period, there was a total of 2080 cases reported to experience an outbreak, including a surge from 2021 to 2022.
+In particular, Toronto health institutions underwent 553 cases in 2020, 394 cases in 2021, and 1133 cases in 2022.
 
 ```{r}
-1 + 1
+library(dplyr)
+
 ```
 
-You can add options to executable code like this
+  It is apparent that Respiratory diseases remain the major outbreak in Toronto healthcare institutions throughout the period.
 
@@
 #| echo: false
-2 * 2
+library(opendatatoronto)
+library(dplyr)
+library(janitor)
+library(tidyverse)
+library(knitr)
+library(RColorBrewer)
+library(kableExtra)
+
+package <- show_package("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
+package
+
+# get all resources for this package
+resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% 'csv')
+
+# load the first datastore resource as a sample
+outbreak_2022 <- filter(datastore_resources, row_number()==15) %>% 
+  get_resource()
+
+# load the
+outbreak_2021 <- filter(datastore_resources, row_number()==14) %>% 
+  get_resource()
+
+# load the
+outbreak_2020 <- filter(datastore_resources, row_number()==13) %>% 
+  get_resource()
+
+outbreak_2020 %>% clean_names()
+outbreak_2021 %>% clean_names()
+outbreak_2022 %>% clean_names()
+
+#|echo:true
+# Number of occurrences in each type of outbreak throughout the years
+  # Count the outbreaks in each year by type 
+  ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
+    mutate(year = 2020) %>% rename(Occurences = "n")
+  ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
+    mutate(year=2021) %>% rename(Occurences = "n")
+  ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak) %>% 
+    mutate(year=2022) %>% rename(Occurences = "n")
+
+  # Merge dataframes
+  ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022)
+
+  # Generate graph
+  ob_by_type$year <- as.factor(ob_by_type$year)
+  ggplot(data=ob_by_type, aes(x=year, y=Occurences, group=Type.of.Outbreak)) +
+    geom_line(aes(color=Type.of.Outbreak), size=0.8)+
+    geom_point(aes(color=Type.of.Outbreak))+
+    scale_linetype_manual(values=c("solid","dotted","twodash"))+
+    scale_color_brewer(palette="Dark2")+
+    theme_minimal()
 ```
 
-The `echo: false` option disables the printing of code (only output is displayed).
+```{r}
+library(dplyr)
+
+```
+
+aaaa
+
+```{r}
+
+```
+
+## Reference

### 2023-02-01 scripts/01_download_data.R

@@ -2,18 +2,15 @@ library(opendatatoronto)
 library(dplyr)
 
 # get package
-package <- show_package("99ff3657-b2e7-4005-a6fd-c36838ccc96d")
-package
+package <- show_package("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
 
 # get all resources for this package
-resources <- list_package_resources("99ff3657-b2e7-4005-a6fd-c36838ccc96d")
+resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
 
 # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% 'csv')
-datastore_resources
+all_data <- filter(resources, tolower(format) %in% 'csv')
 
-# load the immunization coverage in the AY 2017-2018
-immunization_2017_2018 <- filter(datastore_resources, row_number()==1) %>% get_resource()
+# save data set
+write.csv(all_data,"/inputs/data/raw_data.csv")
+raw_data <- read.csv("inputs/data/raw_data.csv")
 
-# load the immunization coverage in the AY 2018-2019
-immunization_2018_2019 <- filter(datastore_resources, row_number()==1) %>% get_resource()

### 2023-02-01 scripts/03_clean_data.R

@@ -1,23 +1,81 @@
 library(opendatatoronto)
 library(dplyr)
 library(janitor)
+library(tidyverse)
+library(knitr)
+library(RColorBrewer)
+library(kableExtra)
+
+
 
 outbreak_2020 %>% clean_names()
 outbreak_2021 %>% clean_names()
 outbreak_2022 %>% clean_names()
 
-ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak)
-ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak)
-ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak)
+# Total number of outbreaks
+  
 
-ob_by_type <- merge(merge(ob_by_type_2020, ob_by_type_2021, by = "Type.of.Outbreak", all.x = TRUE), ob_by_type_2022, by = "Type.of.Outbreak", all.x = TRUE)
 
-ob_by_type_2 <- merge(merge(ob_by_type_2020, ob_by_type_2021, by = 0), ob_by_type_2022, by = 0)
+# Number of occurrences in each type of outbreak throughout the years
+  # Count the outbreaks in each year by type 
+  ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
+    mutate(year = 2020) %>% 
+    rename(Occurences = "n")
+  ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
+    mutate(year=2021)%>% 
+    rename(Occurences = "n")
+  ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak) %>% 
+    mutate(year=2022)%>% 
+    rename(Occurences = "n")
+  
+  # Merge dataframes
+  ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022)
 
-ob_by_type %>% rename("Type.of.Outbreak" = "Outbreak_Type")
+  # Generate graph
+  ob_by_type$year <- as.factor(ob_by_type$year)
+  ggplot(data=ob_by_type, aes(x=year, y=Occurences, group=Type.of.Outbreak)) +
+    geom_line(aes(color=Type.of.Outbreak), size=0.8)+
+    geom_point(aes(color=Type.of.Outbreak))+
+    scale_linetype_manual(values=c("solid","dotted","twodash"))+
[…]

### 2023-02-01 outputs/paper/paper.qmd
Small changes on the graphs position

--- 2023-01-27 outputs/paper/paper.qmd
@@
 ---
-title: "Analysis of Immunization Coverage in Toronto"
+title: "Outbreak Management in Toronto Health Institutions"
 author: "Julie Nguyen"
-format: pdf
-abstract: this is an abstract
+date: "`r format(Sys.time(), '%d %b %Y')`"
+format:
+  pdf:
+    documentclass: scrartcl
+    papersize: a4
+    mainfont: "Garamond"
+    sansfont: "Garamond"
+thanks: 'Code and data are available at: https://github.com/julieenguyennn/Outbreak-Management-in-Toronto-Health-Institutions'
+abstract: Outbreak management has always been one of the most important mission for healthcare institutions, not only to prevent transmissions of diseases/illnesses but also to protect public health.
+Especially in the COVID-19 era, outbreak response and management is considered as the top priority in healthcare institution operation.
+This paper discusses the big picture of outbreak management in Toronto health institutions from 2020 to 2022, as well as pointing out the relationship between healthcare setting, outbreak type and outbreak response.
+number-sections: true
+bibliography: reference.bib
+execute: 
+  message: false
+  warning: false
+  echo: false
 ---
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(echo = FALSE)
 
-## Quarto
+library(opendatatoronto)
+library(dplyr)
+library(janitor)
+library(tidyverse)
+library(knitr)
+library(RColorBrewer)
+library(kableExtra)
+```
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+## Introduction
 
-## Running Code
+  In the previous years, a lot has been learned with regards to the COVID-19 pandemic.
+With the unprecedented cases increase of different diseases, the responsibility for outbreak response and management in healthcare institutions has become more important than ever before.
+Outbreak management helps stop the further transmission of illnesses/diseases by identifying the source and implementing control measures.
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+  In order to protect and improve public health, the Toronto Public Health works with healthcare institutions to prevent and control institutional outbreaks, including the use of best practices in infection prevention control.
+Healthcare institutions are required to monitor staff and patients/residents for signs and symptoms of gastroenteric and respiratory infections under the Ontario Health Protection and Promotion Act (HPPA) (@datatoronto).
+The institutions must also actively seek, detect and report suspected and/or confirmed outbreaks to their local public health unit.
+
+  In the following paper, I will delve into the Outbreaks in Toronto Healthcare Institutions in the previous years to shed light on the differences and trends in reacting to outbreaks.According to opendatatoronto [@datatoronto], an outbreak is defined as "a localized increase (e.g.
+in an institution, or a specific ward or floor within an institution) in the rate of infection or illness, above that which is expected." I will first guide you through the data collection and analyzing process, which will be explained thoroughly in Section 2.
+Then I will visualize the data and explain the implications of the data, finally draw some limitations and conclusions of this data set.
+
+## Data
+### Source
+
+  All the data used in this paper was obtained from the City of Toronto Open Data Portal and is titled "Outbreaks in Toronto Healthcare Institutions".
+The dataset is published by Toronto Public Health and is refreshed weekly, with the lastest refresh on Jan 19, 2023.
+The data provided on the portal include subsets of data from the year 2003 to the time of writing, which is 2023.
+For each year, the dataset displays each observation's details, namely:
+
+  1. Outbreak setting: the type of facility, in terms of the type of healthcare provided
+
+  2. Type of Outbreak: The mode of transmission classification description of the outbreak
+
+  3. Causative & Etiological Agent: the aetiological agent(s) that have been identified from one or more outbreak-related cases
+
+  4. Date Outbreak Began & Date Declared Over
+
+  5. Active Status
+
+  6. Institution Name & Address
+
+### Methodology
+  Based on the dataset, I first generated some questions that this paper aim to answer to gain a better understanding of outbreak management over the COVIC-19 in Toronto health institutions, specifically:
+
+  (1) What type of outbreak was the most encountered by the institutions during the period?
+  (2) Which setting was more prone to expect an outbreak?
+  (3) How was the response rate for each type of outbreak in each setting?
+
+  In order to answer those questions, I chose the 03 dataset that lies in the COVID-19 era - 2020, 2021, and 2022, to analyze, compare, and contrast.
+
+  During the process, all the necessary libraries and packages, tidyverse [@tidyverse], janitor [@janitor], ggplot2 [@ggplot], RColorBrewer [@RColor], and kableExtra [@kableExtra], are utilised.
+
+## Discussion
+   To comprehend the context, I first observed the total cases that were recorded.
+Throughout the period, there was a total of 2080 cases reported to experience an outbreak, with a surge from 2021 to 2022.
+In particular, Toronto health institutions underwent 553 cases in 2020, 394 cases in 2021, and 1133 cases in 2022.
+(@ref(tab:no_outbreak))
 
 ```{r}
-1 + 1
+
+outbreak_2020 <- read.csv(here::here("inputs/data/outbreak_2020.csv"))
+outbreak_2021 <- read.csv(here::here("inputs/data/outbreak_2021.csv"))
+outbreak_2022 <- read.csv(here::here("inputs/data/outbreak_2022.csv"))
 ```
 
-You can add options to executable code like this
+```{r no_outbreak}
+# Total number of outbreaks
+no_outbreak <-
+  data.frame(Year = c("2020","2021","2022"), Outbreak_Amount = c(nrow(outbreak_2020),nrow(outbreak_2021),nrow(outbreak_2022)))
+```
 
 ```{r}
-#| echo: false
-2 * 2
+# Generate graph
+  no_outbreak %>%
+    knitr::kable(caption = "Total number of outbreaks from 2020 - 2022", 
+                 col.names = c("Year", "No. of Outbreak"),
+                 align = c('l', 'l'),
+                 booktabs = T) %>%
+    kable_styling(full_width = FALSE, position = "center", latex_options = c("striped", "HOLD_position"))
 ```
 
-The `echo: false` option disables the printing of code (only output is displayed).
+  Looking deeper into these outbreaks, it is obvious that Respiratory diseases remain the major type of outbreak throughout the period, with a surge of approximately 700 cases from 2021 to 2022.
+While Enteric and other types of outbreak only account for a very small proportion of the outbreaks in total, these two types also remain constant throughout the period.
+See @ for illustration.
+
+```{r}
+outbreak_2020 <- read_csv(here::here("inputs/data/outbreak_2020.csv"))
+outbreak_2021 <- read_csv(here::here("inputs/data/outbreak_2021.csv"))
+outbreak_2022 <- read_csv(here::here("inputs/data/outbreak_2022.csv"))
+
+# Count the outbreaks in each year by type 
+  ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
+    mutate(year = 2020) %>% 
+    rename(Cases = "n")
+  ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
+    mutate(year=2021)%>% 
+    rename(Cases = "n")
+  other_2021 <- data.frame(Type.of.Outbreak = "Other", Cases = 0, year = 2021)
+  ob_by_type_2021 <- bind_rows(ob_by_type_2021, other_2021)
+  ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak) %>% 
+    mutate(year=2022)%>% 
+    rename(Cases = "n")
+
+  # Merge dataframes
+  ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022)
+```
+
+```{r}
+#| label: fig-occurencebyyear
+#| fig-cap: Outbreaks occured in 2020 - 2022 by type
+  # Generate graph
+  ob_by_type$year <- as.factor(ob_by_type$year)
+  ggplot(ob_by_type, aes(fill=Type.of.Outbreak, y=Cases, x=year)) + 
+    geom_bar(position="dodge", stat="identity") +
+    geom_text(
+      aes(label = Cases),
+      colour = "black", size = 3,
+      vjust = -0.2, position = position_dodge(.9)) +
+    guides(fill=guide_legend(title="Type of Outbreak"))
+```
+
+  When examining the correlation between the outbreak type and the setting, it is noticed that the majority of outbreaks happened in Long-term Care Housings (LTCH).
+In 2020, those LTCHs witnessed 20 cases of Enteric outbreak and 296 cases of Respiratory outbreak.
+In 2021, the number of outbreak dropped down to 187 cases in both type, while that of in 2022 increased to more than 450 cases.
+A notable detail to point out is that other types of outbreak only occurred in the Acute Care division of the hospitals.
+In 2022, Toronto also encountered the emergence of outbreak in Transitional Care with 9 cases.
+It is also noticeable that
+
+```{r}
+# Number of outbreaks happened in each setting
+  # Count the outbreaks in each setting by type 
+  ob_setting_2020 <- outbreak_2020 %>% group_by(Type.of.Outbreak) %>% count(Outbreak.Setting) %>% arrange(Type.of.Outbreak, Outbreak.Setting)
+  ob_setting_2021 <- outbreak_2021 %>% group_by(Type.of.Outbreak) %>% count(Outbreak.Setting)%>% arrange(Type.of.Outbreak, Outbreak.Setting)
+  ob_setting_2022 <- outbreak_2022 %>% group_by(Type.of.Outbreak) %>% count(Outbreak.Setting)%>% arrange(Type.of.Outbreak, Outbreak.Setting)
+
+  # Merge dataframes
+  ob_setting <- merge(merge(ob_setting_2020,ob_setting_2021, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE), ob_setting_2022, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE) %>%
+    rename("2020" = n.x, "2021" = n.y, "2022" = n)
+
+  # Generate graph
+  cleaned_ob_setting <- select(ob_setting, c(Outbreak.Setting, "2020", "2021", "2022")) 
+    colnames(cleaned_ob_setting) <- c("","2020","2021","2022")
+  names(cleaned_ob_setting$Outbreak.Setting) <- NULL
+  kbl(cleaned_ob_setting,caption = "The number cases of outbreak by setting from 2020 - 2022", booktabs = TRUE) %>% 
+    kable_styling(full_width = TRUE, latex_options = c("striped", "hold_position")) %>% 
+    column_spec(1, width = "8cm") %>% 
+    pack_rows("Enteric",start_row = 1, end_row = 4) %>% 
+    pack_rows("Other", start_row = 5, end_row = 5) %>% 
+    pack_rows("Respiratory", start_row = 6, end_row = 11) 
+```
+
+aaaa
+
+```{r}
+# Response rate
+  # Year 2020
+  daybegan_2020 <- date_ob_2020 <- as.Date(outbreak_2020$Date.Outbreak.Began)  
+  dayover_2020 <- date_ob_2020 <- as.Date(outbreak_2020$Date.Declared.Over)
+
+  # Year 2021
+  daybegan_2021 <- date_ob_2021 <- as.Date(outbreak_2021$Date.Outbreak.Began)  
+  dayover_2021 <- date_ob_2021 <- as.Date(outbreak_2021$Date.Declared.Over)
+
+  # Year 2022
+  daybegan_2022 <- date_ob_2022 <- as.Date(outbreak_2022$Date.Outbreak.Began)  
+  dayover_2022 <- date_ob_2022 <- as.Date(outbreak_2022$Date.Declared.Over)
+```
+
+```{r}
+  #Calculate response time for each
+  response_rate_2020 <- outbreak_2020 %>% mutate(daybegan_2020,dayover_2020,response_rate = difftime(dayover_2020,daybegan_2020,unit = "days")) #calculate response rate
+  response_rate_2020$response_rate <- as.numeric(response_rate_2020$response_rate)
+  cleaned_response_rate_2020 <- response_rate_2020 %>% select(c(Outbreak.Setting, Type.of.Outbreak,response_rate)) %>%
+    group_by(Outbreak.Setting, Type.of.Outbreak) %>%
+    summarise(response_rate = round(mean(response_rate), digits = 0), .groups = 'drop') %>% 
+    arrange(Type.of.Outbreak)
+
+  response_rate_2021 <- outbreak_2021 %>% mutate(daybegan_2021,dayover_2021,response_rate = difftime(dayover_2021,daybegan_2021,unit = "days"))
+  response_rate_2021$response_rate <- as.numeric(response_rate_2021$response_rate)
+  cleaned_response_rate_2021 <- response_rate_2021 %>% select(c(Outbreak.Setting, Type.of.Outbreak,response_rate)) %>%
+    group_by(Outbreak.Setting, Type.of.Outbreak) %>%
+    summarise(response_rate = round(mean(response_rate), digits = 0), .groups = 'drop') %>% 
+    arrange(Type.of.Outbreak)
+
+  response_rate_2022 <- outbreak_2022 %>% mutate(daybegan_2022,dayover_2022,response_rate = difftime(dayover_2022,daybegan_2022,unit = "days"))
+  response_rate_2022$response_rate <- as.numeric(response_rate_2022$response_rate)
+  cleaned_response_rate_2022 <- response_rate_2022 %>%
+    filter(Active == "N") %>% 
+    select(c(Outbreak.Setting, Type.of.Outbreak,response_rate)) %>%
+    group_by(Outbreak.Setting, Type.of.Outbreak) %>%
+    summarise(response_rate = round(mean(response_rate), digits = 0), .groups = 'drop') %>% 
+    arrange(Type.of.Outbreak)
+```
+
+```{r}
+  # Merge all data
+  all_response_rate <- merge(merge(cleaned_response_rate_2020, cleaned_response_rate_2021, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE), cleaned_response_rate_2022, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE) %>%
+  rename("2020" = response_rate.x, "2021" = response_rate.y, "2022" = response_rate, "Outbreak Type" = Type.of.Outbreak, "Setting" = Outbreak.Setting)
+```
+
+```{r}
+  # Draw table
+  all_response_rate <- select(all_response_rate, c(Setting, "2020", "2021", "2022"))
+  colnames(all_response_rate) <- c("","2020","2021","2022")
+  names(all_response_rate$Outbreak.Setting) <- NULL
+  kbl(all_response_rate,caption = "Response rate by setting (days)", booktabs = TRUE) %>% 
+    kable_styling(full_width = TRUE, latex_options = c("striped", "hold_position")) %>% 
+    column_spec(1, width = "8cm") %>% 
+    pack_rows("Enteric",start_row = 1, end_row = 4) %>% 
+    pack_rows("Other", start_row = 5, end_row = 5) %>% 
+    pack_rows("Respiratory", start_row = 6, end_row = 11)
+```
+
+## Conclusion
+
+## Reference

### 2023-02-02 outputs/paper/paper.qmd
Update discussion

@@
 ---
-title: "Outbreak Management in Toronto Health Institutions"
+title: "An analysis of outbreak management in Toronto Health Institutions from 2020 - 2022"
 author: "Julie Nguyen"
-date: "`r format(Sys.time(), '%d %b %Y')`"
+date: "`r Sys.time()`"
+date-format: "D MMMM YYYY"
 format:
@@
   1. Outbreak setting: the type of facility, in terms of the type of healthcare provided
-
   2. Type of Outbreak: The mode of transmission classification description of the outbreak
-
   3. Causative & Etiological Agent: the aetiological agent(s) that have been identified from one or more outbreak-related cases
-
   4. Date Outbreak Began & Date Declared Over
-
   5. Active Status
-
   6. Institution Name & Address
@@
   In order to answer those questions, I chose the 03 dataset that lies in the COVID-19 era - 2020, 2021, and 2022, to analyze, compare, and contrast.
-
 During the process, all the necessary libraries and packages, tidyverse [@tidyverse], janitor [@janitor], ggplot2 [@ggplot], RColorBrewer [@RColor], and kableExtra [@kableExtra], are utilised.
@@
 In particular, Toronto health institutions underwent 553 cases in 2020, 394 cases in 2021, and 1133 cases in 2022.
-(@ref(tab:no_outbreak))
+(@tbl-totaloutbreak)
 
 ```{r}
-
 outbreak_2020 <- read.csv(here::here("inputs/data/outbreak_2020.csv"))
@@
 ```{r}
+#| label: tbl-totaloutbreak
+#| tbl-cap: "Total number of outbreaks from 2020 - 2022"
 # Generate graph
   no_outbreak %>%
-    knitr::kable(caption = "Total number of outbreaks from 2020 - 2022", 
-                 col.names = c("Year", "No. of Outbreak"),
-                 align = c('l', 'l'),
+    knitr::kable(col.names = c("Year", "No. of Outbreak"),
+                 align = c('l', 'c'),
                  booktabs = T) %>%
-    kable_styling(full_width = FALSE, position = "center", latex_options = c("striped", "HOLD_position"))
+    kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position"))
 ```
@@
 While Enteric and other types of outbreak only account for a very small proportion of the outbreaks in total, these two types also remain constant throughout the period.
-See @ for illustration.
+See @fig-casebyyear for illustration.
 
@@
   ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
-    mutate(year = 2020) %>% 
+    mutate(Year = 2020) %>% 
     rename(Cases = "n")
   ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
-    mutate(year=2021)%>% 
+    mutate(Year=2021)%>% 
     rename(Cases = "n")
-  other_2021 <- data.frame(Type.of.Outbreak = "Other", Cases = 0, year = 2021)
+  other_2021 <- data.frame(Type.of.Outbreak = "Other", Cases = 0, Year = 2021)
   ob_by_type_2021 <- bind_rows(ob_by_type_2021, other_2021)
   ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak) %>% 
-    mutate(year=2022)%>% 
+    mutate(Year=2022)%>% 
     rename(Cases = "n")
@@
 ```{r}
-#| label: fig-occurencebyyear
+#| label: fig-casebyyear
 #| fig-cap: Outbreaks occured in 2020 - 2022 by type
   # Generate graph
-  ob_by_type$year <- as.factor(ob_by_type$year)
-  ggplot(ob_by_type, aes(fill=Type.of.Outbreak, y=Cases, x=year)) + 
+  ob_by_type$Year <- as.factor(ob_by_type$Year)
+  ggplot(ob_by_type, aes(fill=Type.of.Outbreak, y=Cases, x=Year)) + 
     geom_bar(position="dodge", stat="identity") +
@@
 In 2022, Toronto also encountered the emergence of outbreak in Transitional Care with 9 cases.
-It is also noticeable that
+It is also noticeable that the hospital settings ceased to report cases for Enteric type of outbreak from 2021, on the other hand received more cases on respiratory diseases.
+(@tbl-setting)
 
 ```{r}
+#| label: tbl-setting
+#| tbl-cap: "The number cases of outbreak by setting from 2020 - 2022"
+
 # Number of outbreaks happened in each setting
@@
   names(cleaned_ob_setting$Outbreak.Setting) <- NULL
-  kbl(cleaned_ob_setting,caption = "The number cases of outbreak by setting from 2020 - 2022", booktabs = TRUE) %>% 
-    kable_styling(full_width = TRUE, latex_options = c("striped", "hold_position")) %>% 
+  kbl(cleaned_ob_setting, booktabs = TRUE) %>% 
+    kable_styling(full_width = TRUE, latex_options = c("striped", "HOLD_position")) %>% 
     column_spec(1, width = "8cm") %>% 
@@
 
-aaaa
+  Another significant aspect in outbreak response is the numbers that the outbreak is controlled.
+In this case, I calculated the average days that each type of outbreak are controlled by the settings, which is counted by days, as shown in @tbl-responserate.
+A limitation when looking into the duration that outbreak was controlled is that I only chose the observations that were successfully controlled by the end of 2022.
+  In general, the LTCH had the highest response time with an average of 32 days in 2020, 35 days in 2021 followed by Retirement Home with 27 days in 2020 and 31 days in 2021.
 
@@
   dayover_2022 <- date_ob_2022 <- as.Date(outbreak_2022$Date.Declared.Over)
-```
 
-```{r}
   #Calculate response time for each
@@
 ```{r}
+#|label: tbl-responserate
+#|tbl-cap: "Response rate by setting (days)"
   # Merge all data
@@
   rename("2020" = response_rate.x, "2021" = response_rate.y, "2022" = response_rate, "Outbreak Type" = Type.of.Outbreak, "Setting" = Outbreak.Setting)
-```
 
-```{r}
   # Draw table
@@
   names(all_response_rate$Outbreak.Setting) <- NULL
-  kbl(all_response_rate,caption = "Response rate by setting (days)", booktabs = TRUE) %>% 
+  kbl(all_response_rate, booktabs = TRUE) %>% 
     kable_styling(full_width = TRUE, latex_options = c("striped", "hold_position")) %>% 
@@
 ## Conclusion
+Overall,
+
+Through @tbl-setting and @tbl-responserate, I can conclude that one aspect that leads to longer response time is the number of cases reported in that setting.
+However, in 2022, although there were only 03 cases in experiences other types of outbreak but it took an average of 32 days to control them.
+Therefore, the number of cases doesn't account for the amount of time that the outbreak can be controlled.
+It is influenced by other aspects, such as human resources, supplies, environmental conditions, and systems, etc.
+- which was not mentioned in this dataset by @datatoronto.
 

### 2023-02-02 scripts/01_download_data.R

@@ -10,7 +10,17 @@ resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
 # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
 all_data <- filter(resources, tolower(format) %in% 'csv')
 
+ob_source_2020 <- filter(all_data,row_number()==13) %>% get_resource()
+ob_source_2021 <- filter(all_data,row_number()==14) %>% get_resource()
+ob_source_2022 <- filter(all_data,row_number()==15) %>% get_resource()
+
 # save data set
-write.csv(all_data,"/inputs/data/raw_data.csv")
-raw_data <- read.csv("inputs/data/raw_data.csv")
+write.csv(ob_source_2020,"inputs/data/outbreak_2020.csv")
+outbreak_2020 <- read.csv("inputs/data/outbreak_2020.csv")
+
+write.csv(ob_source_2021,"inputs/data/outbreak_2021.csv")
+outbreak_2021 <- read.csv("inputs/data/outbreak_2021.csv")
+
+write.csv(ob_source_2022,"inputs/data/outbreak_2022.csv")
+outbreak_2022 <- read.csv("inputs/data/outbreak_2022.csv")
 

### 2023-02-02 scripts/03_clean_data.R

@@ -6,76 +6,117 @@ library(knitr)
 library(RColorBrewer)
 library(kableExtra)
 
-
-
-outbreak_2020 %>% clean_names()
-outbreak_2021 %>% clean_names()
-outbreak_2022 %>% clean_names()
+outbreak_2020 <- read.csv(here::here("inputs/data/outbreak_2020.csv"))
+outbreak_2021 <- read.csv(here::here("inputs/data/outbreak_2021.csv"))
+outbreak_2022 <- read.csv(here::here("inputs/data/outbreak_2022.csv"))
 
 # Total number of outbreaks
-  
+no_outbreak <-
+  data.frame(Year = c("2020","2021","2022"), Outbreak_Amount = c(nrow(outbreak_2020),nrow(outbreak_2021),nrow(outbreak_2022)))
 
+# Generate graph
+no_outbreak %>%
+  knitr::kable(col.names = c("Year", "No. of Outbreak"),
+               align = c('l', 'c'),
+               booktabs = T) %>%
+  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position"))
 
-# Number of occurrences in each type of outbreak throughout the years
-  # Count the outbreaks in each year by type 
-  ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
-    mutate(year = 2020) %>% 
-    rename(Occurences = "n")
-  ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
-    mutate(year=2021)%>% 
-    rename(Occurences = "n")
-  ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak) %>% 
-    mutate(year=2022)%>% 
-    rename(Occurences = "n")
-  
-  # Merge dataframes
-  ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022)
+# Count the outbreaks in each year by type 
+ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
+  mutate(Year = 2020) %>% 
+  rename(Cases = "n")
+ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
+  mutate(Year=2021)%>% 
+  rename(Cases = "n")
+other_2021 <- data.frame(Type.of.Outbreak = "Other", Cases = 0, Year = 2021)
[…]

### 2023-02-02 scripts/00_simulation.R
generate simulation

@@ -1,2 +1,7 @@
-simulated_data <- tibble("School" = 1:100, "Vaccine 1" = sample(runif(100,min = 0, max = 100), size = 100, replace = TRUE), "Vaccine 2" = sample(runif(100, min = 0, max = 100), size = 100, replace = TRUE))
+simulated_data <-
+  tibble("Setting" = 1:100,
+         "Cases" = floor(runif(100, min=0, max=100)),
+         "Day Began" = sample(seq(as.Date('2020/01/01'), as.Date('2021/01/01'), by="day"), 100, replace = TRUE),
+         "Day Over" = sample(seq(as.Date('2021/01/01'), as.Date('2022/01/01'), by="day"), 100, replace = TRUE))
+
 simulated_data

### 2023-02-03 outputs/paper/paper.qmd
Finalize paper

@@
 ---
-title: "An analysis of outbreak management in Toronto Health Institutions from 2020 - 2022"
+title: "An analysis on outbreak management in Toronto Health Institutions from 2020 - 2022"
 author: "Julie Nguyen"
@@
   pdf:
-    documentclass: scrartcl
+    documentclass: article
     papersize: a4
+    geometry: margin = 1in
     mainfont: "Garamond"
@@
 thanks: 'Code and data are available at: https://github.com/julieenguyennn/Outbreak-Management-in-Toronto-Health-Institutions'
-abstract: Outbreak management has always been one of the most important mission for healthcare institutions, not only to prevent transmissions of diseases/illnesses but also to protect public health.
+abstract: "Outbreak management has always been one of the most important mission for healthcare institutions, not only to prevent transmissions of diseases/illnesses but also to protect public health.
 Especially in the COVID-19 era, outbreak response and management is considered as the top priority in healthcare institution operation.
-This paper discusses the big picture of outbreak management in Toronto health institutions from 2020 to 2022, as well as pointing out the relationship between healthcare setting, outbreak type and outbreak response.
+This paper discusses the outbreak management in Toronto health institutions from 2020 to 2022, as well as pointing out the relationship between healthcare setting, outbreak type and response duration."
 number-sections: true
@@
 ## Introduction
-
   In the previous years, a lot has been learned with regards to the COVID-19 pandemic.
-With the unprecedented cases increase of different diseases, the responsibility for outbreak response and management in healthcare institutions has become more important than ever before.
+With the unprecedented increase in the number of cases of different diseases, the responsibility for outbreak response and management in healthcare institutions has become more important than ever before.
 Outbreak management helps stop the further transmission of illnesses/diseases by identifying the source and implementing control measures.
@@
   In order to protect and improve public health, the Toronto Public Health works with healthcare institutions to prevent and control institutional outbreaks, including the use of best practices in infection prevention control.
-Healthcare institutions are required to monitor staff and patients/residents for signs and symptoms of gastroenteric and respiratory infections under the Ontario Health Protection and Promotion Act (HPPA) (@datatoronto).
+Healthcare institutions are required to monitor staff and patients/residents for signs and symptoms of gastroenteric and respiratory infections under the Ontario Health Protection and Promotion Act (HPPA) [@hppa].
 The institutions must also actively seek, detect and report suspected and/or confirmed outbreaks to their local public health unit.
 
-  In the following paper, I will delve into the Outbreaks in Toronto Healthcare Institutions in the previous years to shed light on the differences and trends in reacting to outbreaks.According to opendatatoronto [@datatoronto], an outbreak is defined as "a localized increase (e.g.
-in an institution, or a specific ward or floor within an institution) in the rate of infection or illness, above that which is expected." I will first guide you through the data collection and analyzing process, which will be explained thoroughly in Section 2.
-Then I will visualize the data and explain the implications of the data, finally draw some limitations and conclusions of this data set.
+  In the following paper, I will delve into the outbreaks in Toronto healthcare institutions during the COVID-19 period to shed light on the differences and trends in reacting to outbreaks.
+According to opendatatoronto [@datatoronto], an outbreak is defined as "a localized increase (e.g.
+in an institution, or a specific ward or floor within an institution) in the rate of infection or illness, above that which is expected." I will first guide you through the data collection, methodology, and ethical concerns, which will be explained thoroughly in Section 2.
+Then I will discuss the data, visualize, and draw conclusions from the acquired data set.
 
@@
 ### Source
-
-  All the data used in this paper was obtained from the City of Toronto Open Data Portal and is titled "Outbreaks in Toronto Healthcare Institutions".
+  All data used in this paper was obtained from the City of Toronto Open Data Portal and is titled "Outbreaks in Toronto Healthcare Institutions".
 The dataset is published by Toronto Public Health and is refreshed weekly, with the lastest refresh on Jan 19, 2023.
@@
 
-  1. Outbreak setting: the type of facility, in terms of the type of healthcare provided
-  2. Type of Outbreak: The mode of transmission classification description of the outbreak
+  1. Outbreak setting: the type of facility, in terms of the type of healthcare provided, including Respiratory, Enteric (Gastroenteric), and Other
+  2. Type of Outbreak: the mode of transmission classification description of the outbreak
   3. Causative & Etiological Agent: the aetiological agent(s) that have been identified from one or more outbreak-related cases
@@
 ### Methodology
-  Based on the dataset, I first generated some questions that this paper aim to answer to gain a better understanding of outbreak management over the COVIC-19 in Toronto health institutions, specifically:
+  Based on the dataset, I first generated some questions that this paper aims to answer to gain a better understanding of outbreak management over the COVID-19 in Toronto health institutions, specifically:
 
@@
   (2) Which setting was more prone to expect an outbreak?
-  (3) How was the response rate for each type of outbreak in each setting?
+  (3) How long did it take for each type of setting control the outbreaks?
 
   In order to answer those questions, I chose the 03 dataset that lies in the COVID-19 era - 2020, 2021, and 2022, to analyze, compare, and contrast.
-During the process, all the necessary libraries and packages, tidyverse [@tidyverse], janitor [@janitor], ggplot2 [@ggplot], RColorBrewer [@RColor], and kableExtra [@kableExtra], are utilised.
+During the process, all the necessary libraries and packages, R package [@rpackage] tidyverse [@tidyverse], janitor [@janitor], ggplot2 [@ggplot], RColorBrewer [@RColor], and kableExtra [@kableExtra], are utilised.
+
+### Ethical Concerns & Limitations
+  Before diving into the discussion, I would like to make some comments on the ethical concerns and limitations when acquiring and analyzing this dataset.
+The major ethical issue of this dataset is the blind spot created from excluding the reasons that caused the outbreaks.
+Particularly, resources such as human labors, systems, finance, workplace environment, patients' condition and medical history, etc.
+contribute to the outbreak management.
+Although the causative and etiological agent of the outbreaks were stated, the lack of those aforementioned information in the dataset may decrease the validity and capacity to compare the relationship between the settings and how long outbreaks were controlled, as well as limit the opportunity to make better suggestions on how to improve outbreak management in those institutions.
 
 ## Discussion
-   To comprehend the context, I first observed the total cases that were recorded.
+   To comprehend the context, let's first observe the overall landscape, of the total cases that were recorded.
 Throughout the period, there was a total of 2080 cases reported to experience an outbreak, with a surge from 2021 to 2022.
@@
   Looking deeper into these outbreaks, it is obvious that Respiratory diseases remain the major type of outbreak throughout the period, with a surge of approximately 700 cases from 2021 to 2022.
-While Enteric and other types of outbreak only account for a very small proportion of the outbreaks in total, these two types also remain constant throughout the period.
+While Enteric and other types of outbreak only accounted for a very small proportion of the outbreaks in total, these two types also remained constant throughout the period.
 See @fig-casebyyear for illustration.
@@
 #| label: fig-casebyyear
-#| fig-cap: Outbreaks occured in 2020 - 2022 by type
+#| fig-cap: "Outbreaks occured by type, from 2020 - 2022"
   # Generate graph
@@
 
-  When examining the correlation between the outbreak type and the setting, it is noticed that the majority of outbreaks happened in Long-term Care Housings (LTCH).
+  When examining the correlation between the outbreak type and the setting, it is noticeable that the majority of outbreaks happened in Long-term Care Housings (LTCH).
 In 2020, those LTCHs witnessed 20 cases of Enteric outbreak and 296 cases of Respiratory outbreak.
 In 2021, the number of outbreak dropped down to 187 cases in both type, while that of in 2022 increased to more than 450 cases.
-A notable detail to point out is that other types of outbreak only occurred in the Acute Care division of the hospitals.
+Another notable detail to point out is that other types of outbreak only occurred in the Acute Care division of the hospitals.
 In 2022, Toronto also encountered the emergence of outbreak in Transitional Care with 9 cases.
-It is also noticeable that the hospital settings ceased to report cases for Enteric type of outbreak from 2021, on the other hand received more cases on respiratory diseases.
+It is also apparent that hospital settings ceased to report cases for Enteric type of outbreak from 2021, on the other hand received more cases on respiratory diseases.
 (@tbl-setting)
@@
   kbl(cleaned_ob_setting, booktabs = TRUE) %>% 
-    kable_styling(full_width = TRUE, latex_options = c("striped", "HOLD_position")) %>% 
+    kable_styling(latex_options = c("striped", "HOLD_position")) %>% 
     column_spec(1, width = "8cm") %>% 
@@
 
-  Another significant aspect in outbreak response is the numbers that the outbreak is controlled.
-In this case, I calculated the average days that each type of outbreak are controlled by the settings, which is counted by days, as shown in @tbl-responserate.
+  Another significant aspect in outbreak response is the duration that the outbreak is controlled.
+In this case, I calculated the average days that each type of outbreak was controlled by the settings, as shown in @tbl-responserate.
 A limitation when looking into the duration that outbreak was controlled is that I only chose the observations that were successfully controlled by the end of 2022.
-  In general, the LTCH had the highest response time with an average of 32 days in 2020, 35 days in 2021 followed by Retirement Home with 27 days in 2020 and 31 days in 2021.
+  In general, LTCHs had the highest response time in Respiratory disease with an average of 32 days in 2020, 35 days in 2021, followed by Retirement Home with 27 days in 2020 and 31 days in 2021.
+For Enteric disease, Retirement Home group remained stable in controlling the outbreaks throughout the period, with an average of 16 days.
+With the emergence of Respiratory disease in Transitional Care establishments in 2022, the average duration of outbreak control was rather high comparing to other kinds of setting.
 
@@
 ```{r}
-#|label: tbl-responserate
-#|tbl-cap: "Response rate by setting (days)"
+#| label: tbl-responserate
+#| tbl-cap: "Outbreak control duration by setting from 2020 - 2022 (days)"
   # Merge all data
@@
   kbl(all_response_rate, booktabs = TRUE) %>% 
-    kable_styling(full_width = TRUE, latex_options = c("striped", "hold_position")) %>% 
+    kable_styling(latex_options = c("striped", "HOLD_position")) %>% 
     column_spec(1, width = "8cm") %>% 
@@
 ## Conclusion
-Overall,
 
-Through @tbl-setting and @tbl-responserate, I can conclude that one aspect that leads to longer response time is the number of cases reported in that setting.
-However, in 2022, although there were only 03 cases in experiences other types of outbreak but it took an average of 32 days to control them.
-Therefore, the number of cases doesn't account for the amount of time that the outbreak can be controlled.
-It is influenced by other aspects, such as human resources, supplies, environmental conditions, and systems, etc.
-- which was not mentioned in this dataset by @datatoronto.
+  In conclusion, during the period of 2020 to 2022, the total number of outbreaks in Toronto health institutions fluctuated and surged to more than 1100 cases by the end of 2022.
+Specifically, Respiratory diseases accounted for the majority of the cases and most cases emerged in Long-term Care Housings.
+
+  Through @tbl-setting and @tbl-responserate, it seems that one aspect that leads to longer response time is the number of cases reported in that setting.
+However, this cannot be concluded by just looking at this dataset.
+In particular, in 2022, although there were only 03 cases in experiences other types of outbreak but it took an average of 32 days to control them.
+In addition, as mentioned before, the lack of information of labor and environmental factors and the difference in the input for the number of cases and the average response days in 2022 affect the preciseness of the results.
+Therefore, the number of cases doesn't account for the amount of time that the outbreak can be controlled since it was influenced by other aspects - which was not mentioned in this dataset by opendatatoronto [@datatoronto].
 

### 2023-02-03 scripts/02_clean_data.R
add preamble

@@ -1,0 +1,132 @@
+#### Preamble ####
+# Purpose: Clean the data downloaded from opendatatoronto
+# Author: Julie Nguyen
+# Date: 2 February 2022
+# Contact: hongan.nguyen@utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - Need to have downloaded the data on opendatatoronto and saved it to inputs/data
+# - Commit and push to github when change anything
+
+library(opendatatoronto)
+library(dplyr)
+library(janitor)
+library(tidyverse)
+library(knitr)
+library(RColorBrewer)
+library(kableExtra)
+
+outbreak_2020 <- read.csv(here::here("inputs/data/outbreak_2020.csv"))
+outbreak_2021 <- read.csv(here::here("inputs/data/outbreak_2021.csv"))
+outbreak_2022 <- read.csv(here::here("inputs/data/outbreak_2022.csv"))
+
+# Total number of outbreaks
+no_outbreak <-
+  data.frame(Year = c("2020","2021","2022"), Outbreak_Amount = c(nrow(outbreak_2020),nrow(outbreak_2021),nrow(outbreak_2022)))
+
+# Generate graph
+no_outbreak %>%
+  knitr::kable(col.names = c("Year", "No. of Outbreak"),
+               align = c('l', 'c'),
+               booktabs = T) %>%
+  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position"))
+
+# Count the outbreaks in each year by type 
+ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
+  mutate(Year = 2020) %>% 
+  rename(Cases = "n")
+ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
+  mutate(Year=2021)%>% 
+  rename(Cases = "n")
+other_2021 <- data.frame(Type.of.Outbreak = "Other", Cases = 0, Year = 2021)
+ob_by_type_2021 <- bind_rows(ob_by_type_2021, other_2021)
+ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak) %>% 
+  mutate(Year=2022)%>% 
+  rename(Cases = "n")
+
+# Merge dataframes
+ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022)
+
+# Generate graph
[…]

### 2023-02-03 scripts/00_simulation.R
update preamble

@@ -1,3 +1,11 @@
+#### Preamble ####
+# Purpose: Simulate the dataset to be solved
+# Author: Julie Nguyen
+# Date: 2 February 2022
+# Contact: hongan.nguyen@utoronto.ca
+# Pre-requisites: 
+# - Have questions to be solved for the data
+
 simulated_data <-
   tibble("Setting" = 1:100,
          "Cases" = floor(runif(100, min=0, max=100)),

### 2023-02-03 scripts/02_clean_data.R

@@ -3,10 +3,8 @@
 # Author: Julie Nguyen
 # Date: 2 February 2022
 # Contact: hongan.nguyen@utoronto.ca
-# License: MIT
 # Pre-requisites: 
 # - Need to have downloaded the data on opendatatoronto and saved it to inputs/data
-# - Commit and push to github when change anything
 
 library(opendatatoronto)
 library(dplyr)

### 2023-02-03 scripts/03_clean_data.R

@@ -1,122 +1,0 @@
-library(opendatatoronto)
-library(dplyr)
-library(janitor)
-library(tidyverse)
-library(knitr)
-library(RColorBrewer)
-library(kableExtra)
-
-outbreak_2020 <- read.csv(here::here("inputs/data/outbreak_2020.csv"))
-outbreak_2021 <- read.csv(here::here("inputs/data/outbreak_2021.csv"))
-outbreak_2022 <- read.csv(here::here("inputs/data/outbreak_2022.csv"))
-
-# Total number of outbreaks
-no_outbreak <-
-  data.frame(Year = c("2020","2021","2022"), Outbreak_Amount = c(nrow(outbreak_2020),nrow(outbreak_2021),nrow(outbreak_2022)))
-
-# Generate graph
-no_outbreak %>%
-  knitr::kable(col.names = c("Year", "No. of Outbreak"),
-               align = c('l', 'c'),
-               booktabs = T) %>%
-  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position"))
-
-# Count the outbreaks in each year by type 
-ob_by_type_2020 <- outbreak_2020 %>% count(Type.of.Outbreak) %>% 
-  mutate(Year = 2020) %>% 
-  rename(Cases = "n")
-ob_by_type_2021 <- outbreak_2021 %>% count(Type.of.Outbreak) %>% 
-  mutate(Year=2021)%>% 
-  rename(Cases = "n")
-other_2021 <- data.frame(Type.of.Outbreak = "Other", Cases = 0, Year = 2021)
-ob_by_type_2021 <- bind_rows(ob_by_type_2021, other_2021)
-ob_by_type_2022 <- outbreak_2022 %>% count(Type.of.Outbreak) %>% 
-  mutate(Year=2022)%>% 
-  rename(Cases = "n")
-
-# Merge dataframes
-ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022)
-
-# Generate graph
-ob_by_type$Year <- as.factor(ob_by_type$Year)
-ggplot(ob_by_type, aes(fill=Type.of.Outbreak, y=Cases, x=Year)) + 
-  geom_bar(position="dodge", stat="identity") +
-  geom_text(
-    aes(label = Cases),
-    colour = "black", size = 3,
-    vjust = -0.2, position = position_dodge(.9)) +
-  guides(fill=guide_legend(title="Type of Outbreak"))
-
[…]

### 2023-02-03 outputs/paper/paper.qmd
update preamble

@@
 Especially in the COVID-19 era, outbreak response and management is considered as the top priority in healthcare institution operation.
-This paper discusses the outbreak management in Toronto health institutions from 2020 to 2022, as well as pointing out the relationship between healthcare setting, outbreak type and response duration."
+This paper discusses the outbreak management in Toronto health institutions from 2020 to 2022, as well as pointing out the relationship between healthcare setting, outbreak type and response duration in order to improve the outbreak management in Toronto health institutions."
 number-sections: true
@@
   In order to answer those questions, I chose the 03 dataset that lies in the COVID-19 era - 2020, 2021, and 2022, to analyze, compare, and contrast.
-During the process, all the necessary libraries and packages, R package [@rpackage] tidyverse [@tidyverse], janitor [@janitor], ggplot2 [@ggplot], RColorBrewer [@RColor], and kableExtra [@kableExtra], are utilised.
+During the process, all the necessary libraries and packages, R package [@rpackage] tidyverse [@tidyverse], janitor [@janitor], ggplot2 [@ggplot], RColorBrewer [@RColor], kableExtra [@kableExtra], and testthat [@testthat] are utilised.
 

### 2023-02-03 scripts/02_clean_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Clean the data downloaded from opendatatoronto
+# Purpose: Clean the data downloaded from opendatatoronto and generate graph
 # Author: Julie Nguyen
 # Date: 2 February 2022
 # Contact: hongan.nguyen@utoronto.ca

### 2023-02-03 tests/testthat.R

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/tests.html
+# * https://testthat.r-lib.org/reference/test_package.html#special-files
+
+library(testthat)
+library(Paper1)
+
+test_check("Paper1")

### 2023-02-03 tests/testthat/04_test.R

@@ -1,0 +1,27 @@
+#### Preamble ####
+# Purpose: Test the code if it's working
+# Author: Julie Nguyen
+# Date: 2 February 2022
+# Contact: hongan.nguyen@utoronto.ca
+# Pre-requisites: 
+# - Generate code
+# - Finish cleaning and generating graphs
+
+test_that("check type of ob_by_type", {
+  expect_s3_class(ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022), "data.frame")
+})
+
+test_that("check if the code generates dataframe type",{
+  expect_s3_class(ob_setting <- merge(merge(ob_setting_2020,ob_setting_2021, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE), ob_setting_2022, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE) %>% 
+                    rename("2020" = n.x, "2021" = n.y, "2022" = n), "data.frame")
+  
+})
+
+test_that("check if the formula generates Date type",{
+  expect_s3_class(daybegan_2020 <- date_ob_2020 <- as.Date(outbreak_2020$Date.Outbreak.Began), "Date")
+})
+
+test_that("check if it can converts from Date to numeric",{
+  expect_type(response_rate_2020$response_rate <- as.numeric(response_rate_2020$response_rate), "numeric")
+})
+

### 2023-02-03 outputs/paper/paper.qmd
finalize and polish paper

@@
 
-  1. Outbreak setting: the type of facility, in terms of the type of healthcare provided, including Respiratory, Enteric (Gastroenteric), and Other
-  2. Type of Outbreak: the mode of transmission classification description of the outbreak
+  1. Outbreak setting: the type of facility, in terms of the type of healthcare provided
+  2. Type of Outbreak: the mode of transmission classification description of the outbreak, including Respiratory, Enteric (Gastroenteric), and Other
   3. Causative & Etiological Agent: the aetiological agent(s) that have been identified from one or more outbreak-related cases
@@
 Another notable detail to point out is that other types of outbreak only occurred in the Acute Care division of the hospitals.
-In 2022, Toronto also encountered the emergence of outbreak in Transitional Care with 9 cases.
+In 2022, Toronto also encountered the emergence of outbreak in Transitional Care with 9 cases, all of which caused by Respiratory disease.
 It is also apparent that hospital settings ceased to report cases for Enteric type of outbreak from 2021, on the other hand received more cases on respiratory diseases.

### 2023-02-03 tests/testthat/04_test.R
organize files

@@ -1,27 +1,0 @@
-#### Preamble ####
-# Purpose: Test the code if it's working
-# Author: Julie Nguyen
-# Date: 2 February 2022
-# Contact: hongan.nguyen@utoronto.ca
-# Pre-requisites: 
-# - Generate code
-# - Finish cleaning and generating graphs
-
-test_that("check type of ob_by_type", {
-  expect_s3_class(ob_by_type <- bind_rows(ob_by_type_2020,ob_by_type_2021,ob_by_type_2022), "data.frame")
-})
-
-test_that("check if the code generates dataframe type",{
-  expect_s3_class(ob_setting <- merge(merge(ob_setting_2020,ob_setting_2021, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE), ob_setting_2022, by = c("Type.of.Outbreak","Outbreak.Setting"), all = TRUE, sort = TRUE) %>% 
-                    rename("2020" = n.x, "2021" = n.y, "2022" = n), "data.frame")
-  
-})
-
-test_that("check if the formula generates Date type",{
-  expect_s3_class(daybegan_2020 <- date_ob_2020 <- as.Date(outbreak_2020$Date.Outbreak.Began), "Date")
-})
-
-test_that("check if it can converts from Date to numeric",{
-  expect_type(response_rate_2020$response_rate <- as.numeric(response_rate_2020$response_rate), "numeric")
-})
-

### 2023-02-03 tests/04_test.R
organize folder

@@ -1,12 +1,0 @@
-# This file is part of the standard setup for testthat.
-# It is recommended that you do not modify it.
-#
-# Where should you do additional test configuration?
-# Learn more about the roles of various files in:
-# * https://r-pkgs.org/tests.html
-# * https://testthat.r-lib.org/reference/test_package.html#special-files
-
-library(testthat)
-library(Paper1)
-
-test_check("Paper1")

### 2023-02-03 scripts/00_simulation.R
explain simulation

@@ -6,10 +6,22 @@
 # Pre-requisites: 
 # - Have questions to be solved for the data
 
-simulated_data <-
-  tibble("Setting" = 1:100,
-         "Cases" = floor(runif(100, min=0, max=100)),
-         "Day Began" = sample(seq(as.Date('2020/01/01'), as.Date('2021/01/01'), by="day"), 100, replace = TRUE),
-         "Day Over" = sample(seq(as.Date('2021/01/01'), as.Date('2022/01/01'), by="day"), 100, replace = TRUE))
+library(tidyverse)
 
-simulated_data
+# Generate a table that portrays the cases in each outbreak type by settings
+
+simulated_data_1 <-
+  tibble("Setting" = sample(c("LTCH","Hospital","Retirement Home","Transitional Care"), 4),
+         "Type" = sample(c("Respiratory", "Enteric"), 4, replace = TRUE),
+         "Cases" = floor(runif(4, min=0, max=100)),
+         )
+simulated_data_1
+
+# Generate a table that has the date outbreak begins and date it's over of each setting by outbreak 
+
+simulated_data_2 <-
+  tibble("Setting" = sample(c("LTCH","Hospital","Retirement Home","Transitional Care"), 4),
+         "Type" = sample(c("Respiratory", "Enteric"), 4, replace = TRUE),
+         "Day Began" = sample(seq(as.Date('2020/01/01'), as.Date('2021/01/01'), by="day"), 4, replace = TRUE),
+         "Day Over" = sample(seq(as.Date('2021/01/01'), as.Date('2022/01/01'), by="day"),4, replace = TRUE))
+simulated_data_2
